### 2024-05-07 PRB - revision reply/reply.tex
revision

\documentclass[12pt]{amsart}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\usepackage[dvipsnames]{xcolor}
\geometry{a4paper} % or letter or a5paper or ... etc
% \geometry{landscape} % rotated page geometry

% See the ``Article customise'' template for come common customisations
% delete this line to display the current date

%%% BEGIN DOCUMENT
\begin{document}
	\color{red} \section{ Report of the First Referee}\\
	The authors compare nuclear spin characterization using ESEEM and DD
	sequences. Overall the work seems timely and appropriate for
	publication in PRB.

	However, I think some sections and figures would benefit from further
	revisions. Below is a list of comments and inquiries:

	1) In some places the authors refer to second order hyperfine
	coupling, I assume this is in regard to the Zeeman splitting of the
	central spin but this could be clarified in the text.

	2) I am not sure whether this is intentional but the equations miss
	punctuation and equation references the brackets.

	3) Some equations are hard to read because they overlap with the text
	of the next column.

	4) Related to 1) additional details to derive the Hamiltonian in Eq.
	(1) would be helpful, for instance when does this equation hold, I
	believe $\Delta \gg$ hyperfine elements is assumed to neglect coupling to Sx and
	Sy and then the nuclear spins are rotated such that there occur no
	terms with Iy? Is there a difference in the derivation for S=1 and
	S=1/2? Maybe an appendix applying the secular approximation and the
	appropriate rotation to the frame leading to this form would be
	helpful for future readers.

	5) Fig. 2 misses the labels for the subfigures the same applies to
	Fig. 3 where I think that additionally the heading of the top left
	panel should read "Electron Spin 1".

	6) In Eq. (2) it would help to provide the state that is used for the
	calculation as well as an intuition why it is sufficient to include
	the exponential decay term after solving the dynamics (and when this
	is valid).

	7) The last paragraph of App. B is very hard to read and would benefit
	from some work on the language.

	8) Could the authors comment on whether they have an intuition whether
	adapted approaches would be feasible for systems with an additional
	central spin like the (15N)V center or defects with even larger spin
	like vanadium defects in silicon carbide?

	9) Could the authors provide a short comment comparing to entanglement
	based approaches for nuclear spin control, e.g.,
	https://www.nature.com/articles/ncomms14660?

	Concluding, if the above points are addressed, I deem the article
	suited for publication within PRB.

	\color{black} \section{Reply to the First Referee}\\
	We appreciate the comments made by the referee. We clarify each point as follows:

	1. Second order hyperfine couplings means terms like $(\frac{A_{zz}}{\omega_L})^2$ and $(\frac{A_{zx}}{\omega_L})^2$. This is stated in the sentence:
	((... the second-order term $\frac{A_{zx}^2}{\omega_L^2}$ can be neglected ... ))

	To clarify this point, we add the following explanation to the text:
	((... \textcolor{blue}{which is the inner product of precession axes of a nuclear spin condition on the electron spin sublevel. This expression indicates the advantage of using DD sequence for sublevels where either $s_0$ or $s_1$ is zero, only first order parallel hyperfine coupling $\frac{A_{zz}}{\omega_L}$ play a role. On the other hand, in spin-1/2 systems ($s_0=-\frac{1}{2}$ and $s_1=\frac{1}{2}$), the first order hyperfine coupling vanishes and only second order parallel hyperfine coupling $(\frac{A_{zz}}{\omega_L})^2$ and second order perpendicular hyperfine coupling $(\frac{A_{zx}}{\omega_L})^2$ remain present. This indicates weak sensitivity of DD sequence on different nuclear spins}...))

	2.

	3.

	4.

	5.

	6.

	7.

	8.

	9.

	\color{red} \section{ Report of the Second Referee}
	The authors propose a blueprint for efficient nuclear spin
	characterization with color centers. The authors claim that they
	propose a more straightforward approach for determining the hyperfine
	interactions among each nuclear and the electron spin. The manuscript
	is written clearly by comparing dynamical decoupling (DD) and Electron
	Spin Echo Envelop Modulation (ESEEM), the numerical results are valid
	and may could be of use in other future experiments. I think the
	results obtained by the authors are of interest for Phys. Rev. B
	readers, so I support the publication in its present form.

	As some minor remarks which the authors may take into account to
	beatify their study further, I would like to mention:

	1) There are no a, b, c and d in Fig. 2, Fig. 3 and Fig.5. There is
	lack of definitions and units of “Hyperfine Parameters” in abscissa in
	Fig. 2 and Fig. 3. This sentence is confusing: “The first 23 hyperfine
	parameters are Azz and the second 23 parameters are Azx of the nuclear
	spin register.” What are the specific meanings of hyperfine parameters
	along horizontal axis and vertical axis in Figs2?

	2) Are there any reasons for the exponential decay part in gray for
	almost all the equations? It is a little bit strange, and the
	exponential decay part looks not in same size as well.

	3) Styles of references [9], [14], [19] and [21] are not uniform with
	other references.

	\color{black} \section{Reply to the Second Referee}

	1.

	2.

	3.

	\color{red} \section{Report of the Third Referee}
	Zahedian et al. present a theoretical study on a the appropriateness of
	various methods of nuclear spin spectroscopy using electron spins in
	solid state materials as probes. They compare dynamical decoupling
	based techniques and consider how using a spin-1/2 system as a probe
	differs to the more common case of a spin-1 (such as the NV center in
	diamond). As the authors note, spin-1/2 electron spin systems are
	emerging in materials such as diamond and silicon and so expanding the
	existing nuclear spectroscopy machinery to this case will be of
	interest to the community.

	The paper and its main result that correlation-style dynamical
	decoupling sequences are most appropriate for performing nuclear spin
	spectroscopy with spin-1/2 electron spin systems should be of interest
	to the community, and so in principle I think it is appropriate to
	publish in PRB. In parts I found the arguments a little unclear and so
	I would recommend some minor revisions before publication. Some
	specific points:

	1) It is not clear to me why the correlation sequence performs much
	worse for the spin-1 probe (comparing Fig 2(c) and 3(c)) and there is
	not much discussion of this point in the text. This result also seems
	to conflict with Fig 5. It also does not seem like the spin-1 probe
	should be worse than spin-1/2 in this case. Also, what are the
	parameters used to generate the plots Fig 3c,d - is it simulating the
	72 pulse DDESEEM?

	2) It would be interesting to know, roughly, how the results of e.g.
	Fig 5 scale with properties such as probe T1 and T2. Does this
	strongly impact the optimal sequence choice? Some spin-1/2 systems may
	have short T1 times as they are able to cross relax with other g=2
	spins present in the crystal (or on the diamond surface, for
	instance).

	3) One goal of mapping e.g. a nuclear spin cluster is to also use the
	central electron spin to selectively manipulate individual nuclear
	spins to use as ancillae or implement quantum information protocols.
	Is it straightforward that the spectral resolution offered by the
	correlation style measurements also allow this sort of control? At
	minimum the extra time overhead from using the longer sequence would
	need to be considered in practice. Could the authors comment on the
	outlook here?

	4) Minor comment on figures: some labels are missing and fig 4 appears
	a bit blurry.

	\color{black} \section{ Reply to the Third Referee}

	1.

	2.

	3.

	4.

\end{document}