### 2024-06-03 PRB - revision reply/references_exampe.tex
References are still to be fixed

\documentclass[12pt]{amsart}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\geometry{a4paper} % or letter or a5paper or ... etc
% \geometry{landscape} % rotated page geometry

% See the ``Article customise'' template for come common customisations

\title{asdf}
\author{sdf}
\date{} % delete this line to display the current date

%%% BEGIN DOCUMENT
\begin{document}

\maketitle
\tableofcontents

\section{111}
\subsection{2323}
\cite{}
\bibliography{references}

\end{document}